### 2020-08-28 paper/supplement-3.Rmd
first pass at supplementary materials for #7

---
title: "Supplement 3"
output: html_document
---

(Content pulled from the google doc)

Rob's idea \[case study section\]:

-   Including a spectrum of worst practice - best practice

    -   Worst

        -   Missing data can be observed in how statistical methods are
            > reported

            -   e.g., missing meters gained

    -   less worse

        -   Missing data reported, but no handling stated

            -   e.g., 17% of data is missing, nothing mentioned about
                > deletion/imputation\
                > [[https://www.sciencedirect.com/science/article/pii/S0167945719301939?casa_token=QfaCiQJaBQMAAAAA:tL3pKZG7s8-umTgU-cVZ53ZuPLchja5OExOEnkohWhjLH4NPGtqeFkSWlZq0EiPG9ip-ZnQ]{.ul}](https://www.sciencedirect.com/science/article/pii/S0167945719301939?casa_token=QfaCiQJaBQMAAAAA:tL3pKZG7s8-umTgU-cVZ53ZuPLchja5OExOEnkohWhjLH4NPGtqeFkSWlZq0EiPG9ip-ZnQ)

            -   e.g., N = 18, and statistical analysis reports N = 14
                > (Robin, tweet, Dave to request paper)

    -   neutral?

        -   There is missing data, and this is reported?

    -   good

        -   Missing data reported, handling is reported, might be bad
            > imputation like mean imputation / data not shared.

    -   best

        -   Missing data reported, handling is reported, appropriate
            > imputation, data shared.

Appendix of techniques for dealing with missing values

-   scripts for exploring missingness

-   How to unpack the missingness in the paper text

```{=html}
<!-- -->
```
-   Ways to visualise missing data for reviewers, and readers of paper
    > to get a feel of how much to take away.

    -   So the meters gained example, I think a vis could be something
        > like the figure below. Also a way to visualise missing values
        > in athlete studies over time, if for whatever reason on
        > X1,x7,x35 day couldn't do a measurement on athlete 1.
        > \^\^instead of google sheet something ggplot\^\^

        -   vis_miss from visdat/naniar:

![](media/image2.png){width="6.267716535433071in"
height="0.8888888888888888in"}

\^\^Background in sports and exercise science\^\^

-   Missing data (or the absence of missing data) is often not reported

-   In general, imputation is not used

-   Missing data is never really discussed

-   Data are usually fit with ANOVA's, therefore values are subbed in,
    > e.g., mean. (really common practice)

**Notes/examples**

Example 1:
[[https://www.sciencedirect.com/science/article/abs/pii/S1440244018303141]{.ul}](https://www.sciencedirect.com/science/article/abs/pii/S1440244018303141)

Problem---The most important predictor, meters gained relative isn\'t
available in the vast majority of the dataset (2001-2008 for example)
(which is where the \'\'prediction accuracy\'\' was the worst. This
isn\'t mentioned anywhere in the paper, that their most important
predictor isn\'t available. So as a reviewer, unless data is provided,
or code shared. You are relying on domain knowledge to know this.

The supplementary materials which provides coefficients for the GLMS
fitted

![](media/image13.png){width="6.270833333333333in"
height="5.888888888888889in"}

No meters gained? This is weird right, talk about a circle, we have
meters gained most important, but GLM perform similar to decision tree
but the GLM doesn't have meters gained (the most important variable)
according to the decision trees?

Oh and meters gained isn't available throughout the dataset.

"GLMs were also created to predict Score Margin in each timeframe and
the prediction accuracies are similar for each time frame but are higher
than the decision tree prediction accuracies."

And whats with figure 2?

![](media/image15.png){width="6.270833333333333in" height="4.375in"}

The end buckets are supposed to be meters gained difference, but we have
a ?

Example 2:
[[https://twitter.com/championdata/status/472174368059957248]{.ul}](https://twitter.com/championdata/status/472174368059957248)

![](media/image5.png){width="3.7031255468066493in"
height="4.275152012248469in"}

![](media/image3.png){width="6.270833333333333in"
height="2.8333333333333335in"}

So if meters gained was first recorded in 2007, what does this mean for
the time frames being compared?

![](media/image11.png){width="6.270833333333333in"
height="2.2222222222222223in"}

Like OK that is 2009-2016, but what about the 2001-2016 was meters
gained only used for 2007-2016 so what was inputted 2001-2007 and if we
compared it to 2009-2016 where meters gained was FULLY available we have
a worse win/loss pred and a in the dataset with the MOST IMPORTANT
VARIABLE AVAILABLE ALL THE WAY THROUGH IT I am really confused.

**Modelling Match Outcome in Australian Football: Improved accuracy with
large databases **

(similar authors as above---Tran, Young etc)

![](media/image10.png){width="3.1875240594925636in"
height="0.7656889763779527in"}

Problem: The most important variable, in 2019 published without
mentioning its availability, nor are results presented in a way that
gives the reader an idea of what was done? Seriously your trying to
predict an outcome and literally the variable doesn\'t EXIST what was
entered in?

Data sharing/code sharing would have made this available to reviewers
and perhaps changed the tone of the papers.

Example 3:\
[[https://shapeamerica.tandfonline.com/doi/abs/10.1080/02640414.2015.1066026\#.XpkfaFMzZhE]{.ul}](https://protect-au.mimecast.com/s/FqW3CZYMqwSk7AkMsKELku?domain=shapeamerica.tandfonline.com)

These results on data available to public (basic stuff like counting of
kicks isn\'t reproducible)---i.e., the table here:

![](media/image9.png){width="3.557292213473316in"
height="2.951379046369204in"}

Means should be available?

Example 4:
[[https://journals.sagepub.com/eprint/Pv3bEcp6EQMnPAa24cuP/full]{.ul}](https://journals.sagepub.com/eprint/Pv3bEcp6EQMnPAa24cuP/full)

Problem:
[[https://twitter.com/RankingSw/status/961502281874092032]{.ul}](https://twitter.com/RankingSw/status/961502281874092032)

![A screenshot of a cell phone Description automatically
generated](media/image7.png){width="1.4457130358705161in"
height="2.085090769903762in"}

![](media/image14.png){width="1.4609733158355205in"
height="1.6510684601924759in"}

So why didn\'t the authors of the paper know this? Was it because the
underlying model (AFLPR) wasn\'t available at the time? etc But
certainly as a reviewer this is a thing someone should know before
reviewing?

Example 5:
[[https://www.tandfonline.com/doi/abs/10.1080/02640414.2015.1066511]{.ul}](https://www.tandfonline.com/doi/abs/10.1080/02640414.2015.1066511)

Figure below seems odd:

![](media/image4.png){width="2.2067705599300087in"
height="1.4527034120734907in"}

-   Progression of science, self-correction and the need for
    > 'self-correction to not be strike against your name'

-   Speak about the content in the papers wanting to push the knowledge,
    > not just get rushed for the purpose of a publication.

Recommendations -- Ask the question: what am I being provided here to
review?

-   Barriers

    -   Data privacy agreements (e.g., NDA), the reviewer would have to
        > sign one, maybe this is something that's not can occur, so
        > what's the minimum.

    -   Betting privacy, players union---you couldn't ever share who the
        > slowest player in the NRL same with the AFL

```{=html}
<!-- -->
```
-   Framed from the reviewer's point of view, needs to be a minimum
    > understanding of what is being done in the paper.

-   Exploratory and missingness (common problem in sports, maybe not
    > able to get access to the player)

```{=html}
<!-- -->
```
-   Minimum requirements: please give the summary stats, i.e., the
    > variable name, description, the dates collect, number of [missing
    > values]{.ul}, etc.

    -   Describe the data in some type of meaningful sense

    -   If there's definitions, over what period does the definition
        > hold for.

-   Idea scenario: Start data, end date.

Example from Twitter message

Dave sent a twitter message about this paper here.

[[https://sci-hub.tw/https://doi.org/10.1080/02640414.2020.1740490]{.ul}](https://sci-hub.tw/https://doi.org/10.1080/02640414.2020.1740490)

From the Paper

This study used commercially available data from the 2015 AFL season
supplied by Champion Data, the official statistics provider for the AFL.
An exemption from ethics review was granted by the institutional ethics
committee. The original data set included all 149,449 possession events
from all matches played during the 2015 AFL season (18 teams, 665
players, 206 regular season and finals series games). A data subset was
created to focus the analysis on regular season matches in order to
explore wholeof-league patterns (i.e., finals matches were omitted as
they are only played by teams in the top 8 end-of-season ladder
positions). This subset was filtered further to include only those
sequences of possession events that i) started with a kick-in following
a behind, and ii) ended with the offensive team losing possession
through a score, stoppage, or turnover. These sequences are referred to
herein as "kick-in possession chains", where a sequence begins with the
player who first receives the ball from a kick-in and excludes the
initial disposal out of the goal square, unless the player starting with
possession "plays on" to himself. Possession chains that involved only
one player or were incomplete due to end-of-quarter time restrictions
were excluded. Chains from 2 drawn matches were also excluded. The final
data set used for analysis comprised a total of 1,720 kick-in possession
chains from 194 games.

Now thankfully we can use fitzRoy to check out how many behinds there
were in 2015 as per the subset.

> fitzRoy::get_afltables_stats(start_date=\"2015-01-01\",
> end_date=\"2016-01-01\")%\>% \#Season chosen
>
> select(Season,Home.score, Away.score, Round, Home.team, Away.team,
> Round, AQ4B, HQ4B)%\>%\# variables needed
>
> filter(!Round %in% c(\"EF\", \"SF\",\"QF\",\"PF\", \"GF\"))%\>%
> \#excluding finals
>
> filter(Home.score != Away.score)%\>% \#excluding draws
>
> distinct()%\>%
>
> mutate(totalbehinds=AQ4B+HQ4B)%\>% \#away behinds + home behinds
>
> summarise(totalbehindsin2015=sum(totalbehinds))
>
> totalbehindsin2015
>
> \<int\>
>
> 1 4275
>
> \>
>
> So if each quarter ended with a invalid observation (didn't result in
> a score, turnover etc) this would be pretty high I'd imagine but help
> illustrate the point
>
> 4\*194
>
> So we have 4275 kick ins from behinds, of which at MOST 194\*4 (776)
> would be excluded say if every quarter had a kick after the siren that
> resulted in a behind.
>
> So that leaves us with 4275-776 total kick ins which is 3499 kick in
> ins total
>
> **So this data suggests that only 1720/3499 or 0.491569 of kick ins
> resulted in a chain? in the study.**
>
> This seems not good.
>
> To put another way,
>
> out of the total behinds or ''opportunities for chains to start'' if
> we take the exclusion rule. **" Possession chains that involved only
> one player or were incomplete due to end-of-quarter time restrictions
> were excluded."**
>
> **Then**
>
> **(4275-1720 )/ 4275 or 0.5976608 of chains are EXCLUDED that seems
> massive!**

![](media/image12.png){width="6.270833333333333in"
height="1.8194444444444444in"}

New Paper: Recommending how to share data in sports science

-   Simplest approach

    -   If you report the summary statistics for each variable

    -   Then recreate the data from that

        -   The min-max and sample evenly

        -   the distribution