### 2020-09-08 paper/supplement-1.Rmd
change examples of missing data in supplement one.

---
title: Supplement 1
output: html_document
---

# missing data examples relevant to sports (MAR, MCAR, MNAR)

There are three broad categorisations describing why data can be 
missing: Missing Completely at Random (MCAR), Missing at Random (MAR), 
and Missing Not at Random (MNAR). These categorisations help describe 
whether the missingness occurs by chance (MCAR), due to some observed 
variable (MAR) or an unobserved variable (MNAR), with MCAR being 
the least harmful, and MAR and MNAR biased. Here, we provide an example 
of data MAR, MCAR and MNAR.

Data MCAR could occur if an Australian Football data collector was taking 
a lunch break during the third quarter of a match, resulting in data not being 
recorded for this time period. Data is MCAR, because the missings are 
not systematically related to other variables.

Data MAR could arise if a weighing scale used to measure players' pre- and 
post-match body mass produced more missing values when placed on soft carpet, 
rather than hard tiles. If the scales produced more missing values because of 
the softer surface, and surface type was a recorded variable, data would be MAR, 
because the missingness is related to some observed data (i.e., surface), but 
not related to the value of the variable with the missing values itself (i.e., body mass). 
(ref: https://stefvanbuuren.name/fimd/sec-MCAR.html).

Data MNAR is when the missings are related to something unobservable. 
For example, data MNAR could arise when, if studying doping, a player 
was to remove their head and body hair, to avoid providing a hair sample for testing
[\[4\]](https://www.zotero.org/google-docs/?4Ildjf). In this case, data
is missing (the hair sample), and this information is not captured by
any other variable that was collected (hair length was not recorded).