### 2020-09-08 paper/supplement-1.Rmd
editing missing day examples, supplement one.

---
title: Supplement 1
output: html_document
---

# Missing data examples relevant to football

There are three broad categorisations describing why data can be 
missing: Missing Completely at Random (MCAR), Missing at Random (MAR), 
and Missing Not at Random (MNAR). These categorisations help describe 
whether the missingness occurs by chance (i.e., MCAR), due to some observed 
variable (i.e., MAR) or an unobserved variable (i.e., MNAR), with MCAR being 
the least harmful, and MAR and MNAR biased. Here, we provide an example 
of data MAR, MCAR and MNAR, relevant to football.

Data could be considered MCAR if an Australian Football data collector was taking 
a lunch break during the third quarter of a match, resulting in data not being 
recorded for this time period. Data is MCAR, because the missings are 
not systematically related to other variables.

Data MAR could arise if players' body mass measurements were taken with
a weighing scale that produced more missing values when placed on soft carpet, rather than 
hard tiles, in the team change room. If surface type was recorded, data would be MAR, 
because the missingness is related the surface, but not related to the value of body mass itself. 
(ref: https://stefvanbuuren.name/fimd/sec-MCAR.html).

**this example still needs work** Data MNAR is when the missings are related to something unobservable. 
For example, data MNAR could arise when, if studying doping, a player 
was to remove their head and body hair, to avoid providing a hair sample for testing 
(https://www.zotero.org/google-docs/?4Ildjf). In this case, data is missing (the hair sample), 
and this information is not captured by any other variable that was collected 
(hair length was not recorded).